---
title: "simpleRegress"
author: "Cleo Ware"
date: "Fall 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


```{r}
age <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(age)
summary.lm(age)
```




```{r}
width <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(width)
summary.lm(width)
```


```{r}
weight <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weight)
summary.lm(weight)
```

```{r}
chest <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chest)
summary.lm(chest)
```


```{r}
headlen <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headlen)
summary.lm(headlen)
```

Here is the order from worst to best based on different variables correlated to the length of a bear by their standard error and their adjusted R- squared 

First, Length~Age: Residual standard error: 7.5 on 52 degrees of freedom, adjusted R- squared: 0.50

Second, Length~Headwth: Residual standard error: 7.3 on 52 degrees of freedom, adjusted R- squared: 0.5321

Third, Length~Weight: Residual standard error: 5.4 on 52 degrees of freedom, adjusted R- squared: 0.7

fourth, Length~Neck: Residual standard error: 5.376 on 52 degrees of freedom, adjusted R- squared: 0.7476

Fifth, Length~Chest: Residual standard error: 4.939 on 52 degrees of freedom, adjusted R- squared: 0.791, adjusted R- squared: 0.7869

Sixth, Length~Headlen: Residual standard error: 4.235 on 52 degrees of freedom, 
adjusted R- squared: 0.8434


## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

